---
title: "Project_Customer_Life_Time_Value"
output: 
  html_document:
    theme: paper
    highlight: tango
    toc: true
    toc_depth: 3
    fig_width: 5
    fig_height: 5
---

#### Importing the data

```{r}
library(gam)
library(dplyr)
library(plyr)
library(randomForest)
library(pROC)
library(MASS)
library(caret)
library(ridge)
# Import starting data
transactional.df <- read.csv("ltv.csv", header=TRUE)
```

#### Variables present in the base data set


```{r}
# Getting all users id
userid.vector <- unique(transactional.df)


classMetrics <- function(score, y, cutoff, 
                         type = c("all", "accuracy", "sensitivity", 
                                  "specificity", "ppv", "npv", "precision", 
                                  "recall")) {
  # This command throws an error if the user specifies a "type" that
  # isn't supported by this function
  type <- match.arg(type, several.ok = TRUE)
  
  # Edit me
  #Getting all the observed values
  observed <- y
  
  #Creating a vector for the predicted values
  predicted <- rep(0,length(score))
  
  #Setting the value for the cut off
  predicted[score >= cutoff] <- 1
  
  #Generating the confidence matrix
  conf.mat <- table(factor(predicted, levels = c(0,1)),factor(observed,levels = c(0,1)))
  
  #Creating a blank data frame
  perf <- data.frame(value=numeric())
  
  #Using i as index for creating the return data frame
  i<-1
  rownames.vector <- NA
  #Getting the True Positive, True Negative, False Positive and False negative value
  TP <- sum(observed==1 & predicted==1)
  TN <- sum(observed==0 & predicted==0)
  FP <- sum(observed==0 & predicted==1)
  FN <- sum(observed==1 & predicted==0)
  
  #Setting value based on option sent
  if(type == "all" || type%in%"accuracy") {
    rownames.vector[i] <- "accuracy"
    i<-i+1
    value <- sum(observed==predicted) / length(observed)
    perf <- rbind(perf,round(value,7))
  }
  if(type == "all" || type%in%"sensitivity") {
    rownames.vector[i] <- "sensitivity"
    i<-i+1
    value <- TP / (TP + FN)
    perf <- rbind(perf,round(value,7))
  }
  if(type == "all" || type%in%"specificity") {
    rownames.vector[i] <- "specificity"
    i<-i+1
    value <- TN / (TN +FP)
    perf <- rbind(perf,round(value,7))
  }
    if(type == "all" || type%in%"ppv") {
    rownames.vector[i] <- "ppv"
    i<-i+1
    value <- TP / (TP +FP)
    perf <- rbind(perf,round(value,7))
    }
  if(type == "all" || type%in%"npv") {
    rownames.vector[i] <- "npv"
    i<-i+1
    value <- TN / (TN +FN)
    perf <- rbind(perf,round(value,7))
  }
  if(type == "all" || type%in%"precision") {
    rownames.vector[i] <- "precision"
    i<-i+1
    value <- TP / (TP +FP)
    perf <- rbind(perf,round(value,7))
  }
    if(type == "all" || type%in%"recall") {
    rownames.vector[i] <- "recall"
    i<-i+1
    value <- TP / (TP +FN)
    perf <- rbind(perf,round(value,7))
    }
  row.names(perf) <- rownames.vector
  colnames(perf) <- "value"
  
  #Returning the list with the required values
  list(conf.mat = conf.mat,
       perf = perf)
  
}

plotClassMetrics <- function(score, y, xvar = NULL, yvar = c("accuracy", "sensitivity", 
                                  "specificity", "ppv", "npv", "precision", 
                                  "recall"),
                             flip.x = FALSE) {
  yvar <- match.arg(yvar)
  
  #If xvar variable is sent or not set 
  if(is.null(xvar)) {
    #Setting the label and the x axis if no x variabkle is chosen
    x.axis <- score[order(score)]
    x.label <- "score"
    #Taking the length as the total number of values in this case
    cutof.list <- seq(from = 0, to = 1, by = 1 / (length(score )-1))
  } else {
    #Setting the label and variable based on user input
    x.label <- xvar
    cutof.list <- seq(from = 0, to = 1, by = 1 / (100))
  }
    # cutof.list <- seq(from = 0, to = 1, by = 1 / (length(score )-1))
    if(!is.null(xvar)) {
      #If x is given start by filling empty values
    x.axis <- rep(NA,length(cutof.list))
    }
    y.axis <- rep(NA,length(cutof.list))
    #Iterate from each value of cutoff
    for(i in 1:length(cutof.list)) {
      if(!is.null(xvar)) {
        # Set the x axis variable
      x.axis[i] <- classMetrics(score, y, cutof.list[i], 
                         xvar)$perf[xvar,]
      }
      #Set the y-axis variable
      y.axis[i] <- classMetrics(score, y, cutof.list[i], 
                         yvar)$perf[yvar,]
      
    }
    #If user selecrts the flip the value
    if(flip.x) {
      x.axis <- 1 - x.axis
      x.label <- paste("1 - ",xvar)
      
    }
    #return the plot
    ggplot(mapping = aes(x = x.axis, y = y.axis)) + geom_line(group = 1) + xlab(x.label) + ylab(yvar)
  
  # Edit me
}

# Function which takes two vectors useful for getting required values
#Input variable 
#   x : Vector from which we have to return the value
#   y : Vector from which we compare the value
#   type : type of comparison if either equal,maximum or minimum of the comparison
get.req.value = function (x, y, type = c("equal","max","min"),equal.value = NA) {
  # This command throws an error if the user specifies a "type" that
  # isn't supported by this function
  type <- match.arg(type, several.ok = TRUE)
  #If parameter given as max calculate return the value of x with max value in y
  if(type == "max") {
    return(x[which.max(y)])
    #If parameter given as min calculate return the value of x with min value in y
  } else if(type == "min") {
    return(x[which.min(y)])
    #If parameter given as equal calculate return the value of x with equal value of the parameter equal.value
  } else if(type=="equal") {
    if(length(which(y==equal.value) > 0))
    return(x[which(y==equal.value)])
    else {
      #Return NA if the equal value not present
      return (NA)
    }
  }
}

summary.df <- ddply(transactional.df, ~ id,
                    summarize,
                    start.date = get.req.value (as.Date(date),status,"equal",equal.value = 0),
                    end.date = get.req.value (as.Date(date),status,"equal",equal.value = 2),
                    num.of.day = as.Date(end.date) - as.Date(start.date),
                    gender = unique(gender),
                    tot.pages = sum(pages) ,
                    tot.onsite = sum(onsite),
                    tot.entered = sum(entered),
                    tot.completed = sum(completed),
                    tot.holiday = sum(holiday),
                    tot.logins = sum(status !=2),
                    user.attritrated = !is.na(end.date),
                    frequency = num.of.day / tot.logins,
                    last.active.login.date = get.req.value(date[status!=2],date[status!=2],"max"),
                    last.comp.order=get.req.value(date[completed!=0],date[completed!=0],"max"),
                     avg.pages.transaction=tot.pages/tot.logins,
                     avg.onsite.transaction=tot.onsite/tot.logins,
                     min.pages=min(pages),
                     max.pages=max(pages),
                     min.onsite=min(onsite),
                     max.onsite=max(onsite),
                     sd.onsite=sd(onsite),
                     sd.pages=sd(pages)
                    )


prediction.date <- '2014-12-31'
#current.month<-month(prediction.date,label=TRUE, abbr=FALSE)

#Setting the number of days for all people who didnt have an end date
summary.df$num.of.day[is.na(summary.df$num.of.day)] <- (as.Date(prediction.date) -  as.Date(summary.df$start.date[which(is.na(summary.df$num.of.day))]))

#Frequency
summary.df$frequency[is.na(summary.df$frequency)] <- (as.Date(prediction.date) -  as.Date(summary.df$start.date[which(is.na(summary.df$frequency))])) / summary.df$tot.logins[which(is.na(summary.df$frequency))]

#### Recency
recency<-vector('numeric')
recency<-
ifelse(is.na(summary.df$end.date),
       as.Date(prediction.date)-as.Date(summary.df$last.active.login.date),
       as.Date(summary.df$end.date)-as.Date(summary.df$last.active.login.date))
summary.df$recency<-recency
#### Monetary
revenue<-vector('numeric')
for(i in 1: nrow(summary.df))
{
if(is.na(summary.df$end.date[i])){
   revenue<-c(revenue, length(seq(from=as.Date(summary.df$start.date[i]), to=as.Date(prediction.date), by='month')))
   }else{
   revenue<-c(revenue, length(seq(from=as.Date(summary.df$start.date[i]), to=as.Date(summary.df$end.date[i]), by='month')))
   }
}
summary.df$revenue<-revenue

### Average number of pages visited per minute
ave.no.pages.per.min<-vector('numeric')
for(i in 1:nrow(summary.df) )
{
  ave.no.pages.per.min[i]<- summary.df$tot.pages[i]/summary.df$tot.onsite[i]
}
summary.df$ave.no.pages.per.min<-ave.no.pages.per.min
## Average number of pages visited per hour
ave.no.pages.per.hour<-vector('numeric')
for(i in 1:nrow(summary.df) )
{
  ave.no.pages.per.hour[i]<- (summary.df$tot.pages[i]/summary.df$tot.onsite[i])*60
}
summary.df$ave.no.pages.per.hour<-ave.no.pages.per.hour
### Ratio of holiday cards to total orders
hol.card.to.total.orders<-vector('numeric')
for(i in 1:nrow(summary.df) )
{
  hol.card.to.total.orders[i]<- summary.df$tot.holiday[i]/summary.df$tot.completed[i]
}
summary.df$hol.card.to.total.orders<-hol.card.to.total.orders
### Ratio of orders completed to orders entered
orders.completed.to.entered<- vector('numeric')
for(i in 1:nrow(summary.df) )
{
  orders.completed.to.entered[i]<- summary.df$tot.completed[i]/summary.df$tot.entered[i]
}
summary.df$orders.completed.to.entered<-orders.completed.to.entered
### Difference in total orders entered and total orders completed
diff.entered.completed<-vector('numeric')
for(i in 1:nrow(summary.df) )
{
  diff.entered.completed[i]<-summary.df$tot.entered[i]-summary.df$tot.completed[i]
}
summary.df$diff.entered.completed<-diff.entered.completed
### Gap between last completed order and end subscription date
diff.comp.endsub<-vector('numeric')
diff.comp.endsub<-
ifelse(is.na(summary.df$end.date),
       as.Date(prediction.date)-as.Date(summary.df$last.comp.order),
       as.Date(summary.df$end.date)-as.Date(summary.df$last.comp.order))
summary.df$diff.comp.endsub<-diff.comp.endsub

```

### last three transactions calculation
```{r}
top<-function(x, num, c1,c2)
  {
sorted<-x[with(x,order(x[,c1],x[,c2],decreasing=T)),]
splits<-split(sorted,sorted[,c1])
df<-lapply(splits,head,num)
do.call(rbind.data.frame,df)
}
first.3.transacions = top(transactional.df,3,1,4)
```

###
```{r}
# transactiondetails<-
# first.3.transacions %>% group_by(id) %>%
# summarize(avg.pages.per.transaction.last3 = mean(pages),avg.onsite.time.per.transaction.last3=mean(onsite),
#           ratio.comp.entered.last3 = sum(completed)/sum(entered),
#           diff.comp.entered.last3 = sum(entered)-sum(completed))


transactiondetails <- ddply(first.3.transacions,"id",summarize,avg.pages.per.transaction.last3=mean(pages), avg.onsite.time.per.transaction.last3=mean(onsite),
          ratio.comp.entered.last3 = sum(completed)/sum(entered),
          diff.comp.entered.last3 = sum(entered)-sum(completed))

summary.df = merge(transactiondetails,summary.df, by="id")

```
###Average gap between completed orders
```{r}
### Average gap between completed orders
#detach(package:plyr)

reqdata<-transactional.df[transactional.df$completed==1, ]
avg.gap.comp=ddply(reqdata, ~id, summarize, .(as.Date(date)), avg.completed.order.gap = mean(diff(as.Date(date))))

avg.gap.comp = subset(avg.gap.comp, select = c("id", "avg.completed.order.gap"))

summary.df = merge(avg.gap.comp,summary.df, by="id")
```
##Average gap between visits 

```{r}
### Average gap between visits
avg.gap.visit = ddply(transactional.df, ~id, summarise, .(as.Date(date)), avg.visits.gap = mean(diff(as.Date(date))))
avg.gap.visit = subset(avg.gap.visit, select = c("id", "avg.visits.gap"))

summary.df = merge(avg.gap.visit,summary.df, by="id")

summary.df$avg.completed.order.gap[is.nan(summary.df$avg.completed.order.gap)] <- max(summary.df$avg.completed.order.gap,na.rm = T)

summary.df$ratio.comp.entered.last3[is.nan(summary.df$ratio.comp.entered.last3)] <- 0
# 
summary.df$avg.visits.gap[is.na(summary.df$avg.visits.gap)] <- max(summary.df$avg.visits.gap,na.rm = T)

#ZZZZZ


summary.df <- summary.df[!is.na(summary.df$sd.onsite),]

summary.df <- summary.df[!is.na(summary.df$sd.pages),]

```

```{r}
#Random Forest Starts
# tree.builder.df <- subset(summary.df, select = c("num.of.day","gender","tot.pages","tot.onsite","tot.entered","tot.completed","tot.holiday","tot.logins","user.attritrated","frequency","recency","revenue","ave.no.pages.per.min","ave.no.pages.per.hour","hol.card.to.total.orders","orders.completed.to.entered","diff.entered.completed","diff.comp.endsub"))

tree.builder.df <- subset(summary.df, select = c("avg.visits.gap","avg.completed.order.gap" 
,"avg.pages.per.transaction.last3","avg.onsite.time.per.transaction.last3","ratio.comp.entered.last3" ,"diff.comp.entered.last3",
"num.of.day","gender","tot.pages", "tot.onsite","tot.entered"
,"tot.completed","tot.holiday","tot.logins","user.attritrated","frequency","avg.pages.transaction" ,"avg.onsite.transaction"#,"min.pages" ,"max.pages","min.onsite"
 , "max.onsite" ,"sd.onsite" , "sd.pages","recency","revenue","ave.no.pages.per.hour","hol.card.to.total.orders", "orders.completed.to.entered" ,"diff.entered.completed","diff.comp.endsub"        ))

col <- colnames(tree.builder.df)[colSums(is.na(tree.builder.df)) > 0]

# tree.builder.df <- subset(summary.df, select = c("recency","frequency","num.of.day","revenue","hol.card.to.total.orders"))

# Randomly select 20% of the data to be held out for model validation
test.indexes <- sample(1:nrow(tree.builder.df), 
                       round(0.2 * nrow(tree.builder.df)))
train.indexes <- setdiff(1:nrow(tree.builder.df), test.indexes)


tree.builder.df.train <- tree.builder.df[train.indexes, ]
tree.builder.df.test <- tree.builder.df[test.indexes,]


attriation.rf <- randomForest(user.attritrated ~., data=tree.builder.df.train, mtry=6, importance=TRUE)

# attriation.rf <- randomForest(user.attritrated ~ recency + frequency + num.of.day+revenue, data=tree.builder.df.train, mtry=4, importance=TRUE)

attriation.predict.rf = predict(attriation.rf,newdata=tree.builder.df.test
                                #,type = "prob"
                                )
y.value <- mapvalues(tree.builder.df.test$user.attritrated,c(TRUE,FALSE),c(1,0))

roc.rf <- roc(as.numeric(y.value),attriation.predict.rf)
plot(roc.rf)

roc.rf$auc


varImpPlot(attriation.rf)



classMetrics(score = attriation.predict.rf,y = y.value, cutoff = 0.56,type = "all")
```

####### Task 2
### Random Forest For Task 2
```{r}
# tree.builder.df <- subset(summary.df, select = c("num.of.day","gender","tot.pages","tot.onsite","tot.entered","tot.completed","tot.holiday","tot.logins","user.attritrated","frequency","recency","revenue","ave.no.pages.per.min","ave.no.pages.per.hour","hol.card.to.total.orders","orders.completed.to.entered","diff.entered.completed","diff.comp.endsub"))
 
tree.builder.df <- subset(summary.df, select = c("avg.visits.gap","avg.completed.order.gap"
,"avg.pages.per.transaction.last3","avg.onsite.time.per.transaction.last3","ratio.comp.entered.last3" ,"diff.comp.entered.last3","gender","frequency","avg.pages.transaction" ,"avg.onsite.transaction"#,"min.pages" ,"max.pages","min.onsite"
, "max.onsite" ,"sd.onsite" , "sd.pages","recency","revenue","ave.no.pages.per.hour","hol.card.to.total.orders", "orders.completed.to.entered" ,"diff.entered.completed","diff.comp.endsub"        ))
 
 
 
# tree.builder.df <- subset(summary.df, select = c("recency","frequency","num.of.day","revenue","hol.card.to.total.orders"))
 
# Randomly select 20% of the data to be held out for model validation
test.indexes <- sample(1:nrow(tree.builder.df),
                       round(0.2 * nrow(tree.builder.df)))
train.indexes <- setdiff(1:nrow(tree.builder.df), test.indexes)
 
 
tree.builder.df.train <- tree.builder.df[train.indexes, ]
tree.builder.df.test <- tree.builder.df[test.indexes,]
 
 
revenue.rf <- randomForest(revenue ~., data=tree.builder.df.train, mtry=6, importance=TRUE)
 
# attriation.rf <- randomForest(user.attritrated ~ recency + frequency + num.of.day+revenue, data=tree.builder.df.train, mtry=4, importance=TRUE)
 
predict.rf = predict(revenue.rf,newdata=tree.builder.df.test
                                #,type = "prob"
                                )
y.value <- mapvalues(tree.builder.df.test$user.attritrated,c(TRUE,FALSE),c(1,0))
 
roc.rf <- roc(as.numeric(y.value),attriation.predict.rf)
plot(roc.rf)
 
roc.rf$auc
 
 
varImpPlot(revenue.rf)
 
classMetrics(score = attriation.predict.rf,y = y.value, cutoff = 0.56,type = "all")

mean((tree.builder.df.test$revenue - predict(revenue.rf, tree.builder.df.test))^2)

```

### GAM model for 2nd question
```{r}
summ.df = summary.df
summ.df<-data.frame(lapply(summary.df, function(x) as.numeric(as.character(x))))
summ.df<-summ.df[,c(31,37,2,36,22,19,30)]

listofpredictors<-names(summ.df)
#Split data into test(30%) and train(70%)
n = nrow(summ.df)
trainIndex = sample(1:n, size = round(0.7*n), replace=FALSE)
train = summ.df[trainIndex ,]
test = summ.df[-trainIndex ,]

#gam.formula <- formula(paste("user.attritrated ~ ", paste("poly(",listofpredictors[4,5,6], ", 3)", sep = "", collapse= " + ")))
attrition.gam <- gam(revenue~s(frequency) +s(recency)+ s(avg.pages.transaction)+s(diff.entered.completed)+ s(avg.visits.gap)+s(diff.comp.endsub), data = train, df=5)

par(mfrow = c(15,4))
plot(attrition.gam, col = "forestgreen", lwd=3)
mean((test$revenue - predict(attrition.gam, test))^2)

```

### LDA for classification in 1st Question
```{r}
lda.fit=MASS::lda(user.attritrated~., data=train, na.action=TRUE, CV=TRUE)
conf.mat.lda = table(test$user.attritrated, fit$class)
sum(diag(prop.table(conf.mat.lda)))
```


### KNN model for 2nd question
```{r}
#knn.fit = FNN::knn.reg(train, test=NULL, revenue, k=9, algorithm = c("kd_tree", "cover_tree", "brute"))
knn.fit1 = caret::knnreg(train, train$revenue, k=4, na.action="pass")
plot(test$revenue, predict(knn.fit1, test))

mean((test$revenue - predict(knn.fit1, test))^2)
```


### Ridge Regression model for 2nd question
```{r}
ridge.fit = ridge::linearRidge(revenue~., data = train)
mean((test$revenue - predict(ridge.fit, test))^2)
```

### Pair-Pair plot for highly significant predictors from RF for # 1st question
```{r}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = pmax(1, cex.cor * r))
}
 
pairs(summary.df[,listofpredictors], lower.panel = panel.cor)

```


### Cross-validation check for different models
```{r}
options(warn=-1)  # To suppress warning messages
smoothCV <- function(x, y, K = 10, df.min = 1, df.max = 10) {

  set.seed(1)
  data.stream <- data.frame(x, y)
    mse.error.poly = 0
    mse.error.cubic = 0
    mse.error.smoothing = 0
  
  new.data.stream = data.stream[sample(nrow(data.stream)),]
  data.rows = nrow(new.data.stream)
  folds <- cut(seq(1, data.rows),breaks = K,labels = FALSE)
  
  dframe = NULL

  for (df in df.min:df.max)
  {
    poly.fit = NULL
    cubic.fit = NULL
    smoothingspline.fit = NULL
    mse.error.poly = 0
    mse.error.cubic = 0
    mse.error.smoothing = 0

    for(j in 1:K) 
    {
    
    index = which(folds == j, arr.ind = TRUE)
    testData = new.data.stream[index, ]
    trainData = new.data.stream[-index, ]  
      
    # MSE calculation for cubic spline fit for df greater than 2
    if(df > 2)
    {
    cubic.fit =lm(y ~ bs(x,df=df), data = trainData) 
    predict.cubic = predict(cubic.fit, testData)
    mse.cubic = mean((testData$y - predict.cubic)^2) * nrow(testData) / nrow(new.data.stream)
    mse.error.cubic =  mse.error.cubic + mse.cubic
    } 
    else
    {
     mse.error.cubic = NA 
    }
     
    poly.fit = lm(y ~ poly(x, df), data = trainData)
    predict.poly = predict(poly.fit, testData)
    mse.poly = mean((testData$y - predict.poly)^2) * nrow(testData) / nrow(new.data.stream)
    mse.error.poly = mse.error.poly + mse.poly
    
    if(df > 1)
    {
      smoothingspline.fit = smooth.spline(trainData$x,trainData$y, df = df) 
      predict.smoothing = predict(smoothingspline.fit,testData$x)$y
      mse.smooth = mean((testData$y - predict.smoothing)^2) * nrow(testData) / nrow(new.data.stream)
      mse.error.smoothing =  mse.error.smoothing + mse.smooth
    }
    else
    {
      mse.error.smoothing = NA   
    }
    }

    dframe = rbind(dframe, data.frame(df = df, method = "poly", cv.error = mean(mse.error.poly)))
    dframe = rbind(dframe, data.frame(df = df, method = "cubic.spline", cv.error = mean(mse.error.cubic)))
    dframe = rbind(dframe, data.frame(df = df, method = "smoothing.spline", cv.error = mean(mse.error.smoothing)))

  } 
   dframe 
}

smoothCV(summary.df$tot.holiday, summary.df$user.attritrated, K = 10, df.min = 1, df.max = 10)

```